Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-20 MDPI_submission.Rmd
Commit 2023-12-20 16:09

abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, energy production and can affect all living organisms. 
  Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales. Periods of positive trends are referred as brightening periods and of negative trends as dimming periods. 
  In this study we have used 30 years of pyranometer data in Thessaloniki, Greece, to investigate the variability of SDR under three types of sky conditions (clear-, cloudy- and all-sky). 
  The clear-sky data were identified by applying a cloud screening algorithm. 
  We have found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $\sim 0.1\,\%/\text{year}$ for clear-sky, and $0.41\,\%/\text{year}$ for cloudy conditions.
  The consistency of these trends, their seasonal variability, and the effect of the solar zenith angle have also been investigated.
  Under all three sky categories the SDR trend is stronger in winter with $0.7$, $0.4$ and $0.76\,\%/\text{year}$, respectively for all-, clear- and cloudy-skies. In automn, there are the next larger trends of $0.42$ and $0.19\,\%/\text{year}$, for all- and cloudy-skies respectively. 
  The rest seasonal trends are significant smaller, close to zero, with clear- and cloudy-skies having a negative trend for summer.
  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
  Finally, we are discussing shorter-term variations in SDR anomalies by examining the patterns of the cumulative sums of monthly anomalies from the climatological mean, both before and after removing the long-term trend.

@@

abbreviations:
  - short: DNI
    long: Direct beam/normal irradiance
  - short: CSid
    long: Clear sky identification algorithm
  - short: CUSUM
    long: Cumulative sum
  - short: SDR
    long: Solar downward radiation
  - short: SZA
    long: Solar zenith angle

@@

To calculate the reference clear-sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945 (Eq.\ \ref{eq:hau}), adjusted for our site:
<!-- adjusted Haurwitz -->
\begin{equation}
\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
\end{equation}
where $\theta$ is the SZA.

@@

In the final dataset
<!-- $`r SDR_N_All`$ -->
$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
of the 1-minute data were identified as under clear-sky conditions and
$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
as under cloud-sky conditions.
<!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->

@@

We have analyzed a 30-year dataset of SRD measurements in Thessaloniki Greece 
(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
solar irradiance under different sky conditions.
Under all-sky conditions 
there is a positive trend in SDR of $`r alltrend`\,\%/y$ (brightening). 
A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend of $0.33\,\%/y$.
The slight increase of this trend indicates that the brightening of SDR continues and is likely caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
A smaller trend has
been found under clear-sky conditions ($`r cleartrend`\,\%/y$) which supports the
notion that part of the brightening is caused by decreasing aerosols.
@Siomos2020 have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018.
The attenuation
of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
Unfortunately, for this study aerosol data for the entire period were
not available in order to quantify their effect on SDR. 
The brightening effect on SDR under cloudy-sky conditions
($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is also decreasing
during this period.
As long term data of cloud optical thickness are
also not available for the region, we cannot draw quantitative
conclusions.
<!-- @Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.).  -->
 <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W    ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.)-->
<!-- We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$. -->

>>>>

abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, energy production and can affect all living organisms. 
  Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales. Periods of positive trends are referred as brightening periods and of negative trends as dimming periods. 
  In this study we have used 30 years of pyranometer data in Thessaloniki, Greece, to investigate the variability of SDR under three types of sky conditions (clear-, cloudy- and all-sky). 
  The clear-sky data were identified by applying a cloud screening algorithm. 
  We have found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $\sim 0.1\,\%/\text{year}$ for clear-sky, and $0.41\,\%/\text{year}$ for cloudy conditions.
  The consistency of these trends, their seasonal variability, and the effect of the solar zenith angle have also been investigated.
  Under all three sky categories the SDR trend is stronger in winter with $0.7$, $0.4$ and $0.76\,\%/\text{year}$, respectively for all-, clear- and cloudy-skies. 
  The next larger seasonal trends are in automn $0.42$ and $0.19\,\%/\text{year}$, for all- and cloudy-skies respectively. 
  The rest seasonal trends are significant smaller, close to zero, with a negative values in summer, for clear- and cloudy-skies.
  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
  Finally, we are discussing shorter-term variations in SDR anomalies by examining the patterns of the cumulative sums of monthly anomalies from the climatological mean, both before and after removing the long-term trend.

@@

abbreviations:
  - short: DNI
    long: Direct beam/normal irradiance
  - short: ERA5
    long: ECMWF Reanalysis v5
  - short: CSid
    long: Clear sky identification algorithm
  - short: CUSUM
    long: Cumulative sum
  - short: SDR
    long: Solar downward radiation
  - short: SZA
    long: Solar zenith angle

@@

To calculate the reference clear-sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945 (Eq.\ \ref{eq:hau}), adjusted for our site:
<!-- adjusted Haurwitz -->
\begin{equation}
\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
\end{equation}
where $\theta$ is the SZA.

@@

In the final dataset
<!-- $`r SDR_N_All`$ -->
<!-- $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ -->
$26\,\%$ of the days were identified as under clear-sky conditions and
<!-- $`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$ -->
$48\,\%$ as under cloud-sky conditions.
The remaining $26\,\%$ of the data correspond to mixed cases and were not analyzed as a separate group.
<!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->

@@

We have analyzed a 30-year dataset of SRD measurements in Thessaloniki Greece 
(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
solar irradiance under different sky conditions.
Under all-sky conditions 
there is a positive trend in SDR of $`r alltrend`\,\%/y$ (brightening). 
A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend of $0.33\,\%/y$.
The slight increase of this trend indicates that the brightening of SDR continues and is likely caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
A smaller trend has
been found under clear-sky conditions ($`r cleartrend`\,\%/y$) which supports the
notion that part of the brightening is caused by decreasing aerosols.
@Siomos2020 have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018.
The attenuation
of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
Unfortunately, for this study aerosol data for the entire period were
not available in order to quantify their effect on SDR. 
The brightening effect on SDR under cloudy-sky conditions
($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is also decreasing
during this period.
As long term data of cloud optical thickness are
also not available for the region, we cannot draw quantitative
conclusions.
<!-- @Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.).  -->
 <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W    ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.)-->
<!-- We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$. -->

### 2023-12-20 MDPI_submission.tex
Commit 2023-12-20 16:09

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, energy production and can affect
all living organisms. Observations of SDR at different locations around
the world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred as brightening periods
and of negative trends as dimming periods. In this study we have used 30
years of pyranometer data in Thessaloniki, Greece, to investigate the
variability of SDR under three types of sky conditions (clear-, cloudy-
and all-sky). The clear-sky data were identified by applying a cloud
screening algorithm. We have found a positive trend of
\(0.38\,\%/\text{year}\) for all-sky, \(\sim 0.1\,\%/\text{year}\) for
clear-sky, and \(0.41\,\%/\text{year}\) for cloudy conditions. The
consistency of these trends, their seasonal variability, and the effect
of the solar zenith angle have also been investigated. Under all three
sky categories the SDR trend is stronger in winter with \(0.7\), \(0.4\)
and \(0.76\,\%/\text{year}\), respectively for all-, clear- and
cloudy-skies. In automn, there are the next larger trends of \(0.42\)
and \(0.19\,\%/\text{year}\), for all- and cloudy-skies respectively.
The rest seasonal trends are significant smaller, close to zero, with
clear- and cloudy-skies having a negative trend for summer. The SDR
trend is increasing with increasing solar zenith angle, except under
cloudy skies where the trend is highly variable and close to zero.
Finally, we are discussing shorter-term variations in SDR anomalies by
examining the patterns of the cumulative sums of monthly anomalies from
the climatological mean, both before and after removing the long-term
trend.}

@@

To calculate the reference clear-sky \(\text{SDR}_\text{CSref}\) we used
the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of
\citet{Haurwitz1945} (Eq.~\ref{eq:hau}), adjusted for our site:
\begin{equation}
\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
\end{equation} where \(\theta\) is the SZA.

@@

In the final dataset \(23.3\%\) of the 1-minute data were identified as
under clear-sky conditions and \(43\%\) as under cloud-sky conditions.

@@

We have analyzed a 30-year dataset of SRD measurements in Thessaloniki
Greece (1993 -- 2023) aiming to identify the long term variability of
solar irradiance under different sky conditions. Under all-sky
conditions there is a positive trend in SDR of \(0.38\,\%/y\)
(brightening). A previous study \citep{Bais2013} for the period 1993 --
2011 reported also a positive trend of \(0.33\,\%/y\). The slight
increase of this trend indicates that the brightening of SDR continues
and is likely caused by continuing decreases in aerosol optical depth
and the optical thickness of clouds over the area. A smaller trend has
been found under clear-sky conditions (\(0.097\,\%/y\)) which supports
the notion that part of the brightening is caused by decreasing
aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018. The
attenuation of SDR by aerosols over Europe have been proposed as major
factor by \citet{Wild2021}. Unfortunately, for this study aerosol data
for the entire period were not available in order to quantify their
effect on SDR. The brightening effect on SDR under cloudy-sky conditions
(\(0.41\,\%/y\)), suggests that cloud optical thickness is also
decreasing during this period. As long term data of cloud optical
thickness are also not available for the region, we cannot draw
quantitative conclusions.

@@

\noindent
\begin{tabular}{@{}ll}
DNI & Direct beam/normal irradiance \\
CSid & Clear sky identification algorithm \\
CUSUM & Cumulative sum \\
SDR & Solar downward radiation \\
SZA & Solar zenith angle \\
\end{tabular}}

>>>>

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, energy production and can affect
all living organisms. Observations of SDR at different locations around
the world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred as brightening periods
and of negative trends as dimming periods. In this study we have used 30
years of pyranometer data in Thessaloniki, Greece, to investigate the
variability of SDR under three types of sky conditions (clear-, cloudy-
and all-sky). The clear-sky data were identified by applying a cloud
screening algorithm. We have found a positive trend of
\(0.38\,\%/\text{year}\) for all-sky, \(\sim 0.1\,\%/\text{year}\) for
clear-sky, and \(0.41\,\%/\text{year}\) for cloudy conditions. The
consistency of these trends, their seasonal variability, and the effect
of the solar zenith angle have also been investigated. Under all three
sky categories the SDR trend is stronger in winter with \(0.7\), \(0.4\)
and \(0.76\,\%/\text{year}\), respectively for all-, clear- and
cloudy-skies. The next larger seasonal trends are in automn \(0.42\) and
\(0.19\,\%/\text{year}\), for all- and cloudy-skies respectively. The
rest seasonal trends are significant smaller, close to zero, with a
negative values in summer, for clear- and cloudy-skies. The SDR trend is
increasing with increasing solar zenith angle, except under cloudy skies
where the trend is highly variable and close to zero. Finally, we are
discussing shorter-term variations in SDR anomalies by examining the
patterns of the cumulative sums of monthly anomalies from the
climatological mean, both before and after removing the long-term
trend.}

@@

To calculate the reference clear-sky \(\text{SDR}_\text{CSref}\) we used
the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of
\citet{Haurwitz1945} (Eq.~\ref{eq:hau}), adjusted for our site:
\begin{equation}
\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
\end{equation} where \(\theta\) is the SZA.

@@

In the final dataset \(26\,\%\) of the days were identified as under
clear-sky conditions and \(48\,\%\) as under cloud-sky conditions. The
remaining \(26\,\%\) of the data correspond to mixed cases and were not
analyzed as a separate group.

@@

We have analyzed a 30-year dataset of SRD measurements in Thessaloniki
Greece (1993 -- 2023) aiming to identify the long term variability of
solar irradiance under different sky conditions. Under all-sky
conditions there is a positive trend in SDR of \(0.38\,\%/y\)
(brightening). A previous study \citep{Bais2013} for the period 1993 --
2011 reported also a positive trend of \(0.33\,\%/y\). The slight
increase of this trend indicates that the brightening of SDR continues
and is likely caused by continuing decreases in aerosol optical depth
and the optical thickness of clouds over the area. A smaller trend has
been found under clear-sky conditions (\(0.097\,\%/y\)) which supports
the notion that part of the brightening is caused by decreasing
aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018. The
attenuation of SDR by aerosols over Europe has been proposed as major
factor by \citet{Wild2021}. Unfortunately, for this study aerosol data
for the entire period were not available in order to quantify their
effect on SDR. The brightening effect on SDR under cloudy-sky conditions
(\(0.41\,\%/y\)), suggests that cloud optical thickness is also
decreasing during this period. As long term data of cloud optical
thickness are also not available for the region, we cannot draw
quantitative conclusions.

@@

\noindent
\begin{tabular}{@{}ll}
DNI & Direct beam/normal irradiance \\
ERA5 & ECMWF Reanalysis v5 \\
CSid & Clear sky identification algorithm \\
CUSUM & Cumulative sum \\
SDR & Solar downward radiation \\
SZA & Solar zenith angle \\
\end{tabular}}
